# Methods

## R packages (NonCompart and ncar)

Two R packages (NonCompart and ncar) for NCA were developed in the open-source R programming language in order to allow free public use. R packages can be installed and loaded using the following scripts.

```r
install.packages(c('NonCompart', 'ncar'))
library(NonCompart)
library(ncar)
```

Detailed documentation and examples for each package can be found on the online user manual in the CRAN repository (<http://cran.r-project.org/web/packages/NonCompart/index.html>, <http://cran.r-project.org/web/packages/ncar/index.html>) or directly within the R console by entering ?function (e.g.?NonCompart,?ncar ). These two packages are implemented in R and can accept a set of input arguments that allow for generation of NCA output. The names of most NCA metrics estimated by the function of these packages are consistent with those used in WinNonlin®. (Table \@ref(tab:table1))

```{r table1, echo = FALSE}
kable(read_csv('data-raw/table-1.csv'), caption = 'Description of PK parameters of WinNonlin and the R packages')
```


## Software

WinNonlin® (Pharsight, Mountain View, CA, USA)[4] in Microsoft-Windows 7 (64 bit) was used for the computation. R 3.4.2 in Microsoft-Windows 7 (64 bit) was used for the comparison of calculated values.

## Dataset

To compare the outputs generated by R packages and WinNonlin®, we used Theoph dataset obtained from the R software. The Theoph dataset has 132 observations from 12 subjects. A portion of the Theoph dataset (subject ID = 8) is shown in Table \@ref(tab:table2).

```{r table2, echo = FALSE}
kable(read_csv('data-raw/table-2.csv'), caption = 'Part of Theoph dataset with information on key (subject),
time, and concentration')
```

