# (APPENDIX) Appendix {-}

# Environment

```{r, echo = FALSE}
pkgInfo <- lapply(pkg, packageDescription, fields = c("Package", "Version"))
kable(tibble(Package=sapply(pkgInfo, `[[`, 1), Version=sapply(pkgInfo, `[[`, 2)), booktabs = TRUE)
# devtools::session_info()
```

# References

```{r, include = FALSE}
knitr::write_bib(file = 'bib/packages.bib')
```

1. Acharya C, Hooker AC, Turkyilmaz GY, Jonsson S, Karlsson MO. A diag-
nostic tool for population models using non-compartmental analysis: The
ncappc package for R. Comput Methods Programs Biomed 2016;127:
83-93. doi: 10.1016/j.cmpb.2016.01.013.
2. Gabrielsson J, Weiner D. Non-compartmental analysis. Methods Mol Biol
2012;929:377–389.
3. The R Project for Statistical Computing. R. http://www.r-project.org/
Accessed 6 November 2017.
4. Kinetica. http://www.kinetica.com/ Accessed 28 February 2018
5. Certara. Phenix WinNonlin®. https://www.certara.com/software/pkpd-
modeling-and-simulation/phoenix-winnonlin/?ap%5B0%5D=PKPD/
Accessed 28 February 2018.
6. Kim MG, Yim DS, Bae KS. R-based reproduction of the estimation process
hidden behind NONMEN Part 1: first-order approximation method. Transl
Clin Pharmacol 2015;23:1-7.
7. Study Data Tabulation Model Implementation Guide: Human Clinical Trials
Version 3.2, Clinical Data Interchange Standards Consortium. https://www.
cdisc.org/system/files/all/standard_category/application/pdf/sdtmig_
v3.2.pdf. Accessed 28 February 2018.

